---
title: "Iris Dataset Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(plotly)
library(ggplot2)
library(shiny)

data(iris)
```


```{r}

set.seed(123) # For reproducibility
iris_kmeans <- kmeans(iris[, -5], centers = 3) 

iris$Cluster <- as.factor(iris_kmeans$cluster)




iris_pca <- prcomp(iris[, 1:4])$x[, 1:2] 
iris$PCA1 <- iris_pca[, 1]
iris$PCA2 <- iris_pca[, 2]

```

```{r}

```

# Sidebar {.sidebar}
### About the Iris Dataset

The Iris dataset, also known as Fisher's Iris dataset, was introduced by statistician Ronald Fisher in 1936. It includes measurements from three species of Iris flowers:

- **Iris setosa**
- **Iris virginica** 
- **Iris versicolor**

#### Features Measured:
- Sepal Length (cm)
- Sepal Width (cm)
- Petal Length (cm)
- Petal Width (cm)

#### Dataset Properties:
- 150 total samples
- 50 samples per species
- 4 numeric features
- No missing values


# Summary 
## Row 1
```{r}

plot_ly(data = iris, x = ~PCA1, y = ~PCA2, color = ~Cluster,
        type = "scatter", mode = "markers") %>%
  layout(title = "PCA Plot: Cluster Visualization",
         xaxis = list(title = "PCA1"),
         yaxis = list(title = "PCA2"))


```
<br>


```{r}

iris$Petal.Area <- iris$Petal.Length * iris$Petal.Width
iris$Sepal.Area <- iris$Sepal.Length * iris$Sepal.Width


plot_ly(data = iris, x = ~Sepal.Area, y = ~Petal.Area, color = ~Species,
        type = 'scatter', mode = 'markers', opacity = 0.8) %>%
  layout(title = "Petal Area vs. Sepal Area",
         xaxis = list(title = "Sepal Area (Length × Width)"),
         yaxis = list(title = "Petal Area (Length × Width)"))
```

# Individual Analysis

## Scatter Plot {.tabset}

### Petal

```{r}
plot_ly(data = iris, x = ~Petal.Length, y = ~Petal.Width, color = ~Species,
        type = 'scatter', mode = 'markers')|>
  layout(title = "Petal Dimensions: Length vs. Width",
         xaxis = list(title = "Petal Length"),
         yaxis = list(title = "Petal Width"))

```

### Sepal

```{r}
plot_ly(data = iris, x = ~Sepal.Length, y = ~Sepal.Width, color = ~Species,
        type = 'scatter', mode = 'markers') %>%
  layout(title = "Sepal Dimensions: Length vs. Width",
         xaxis = list(title = "Sepal Length"),
         yaxis = list(title = "Sepal Width"))


```

## Histogram {.tabset}

### Petal Length

```{r}
plot_ly(data = iris, x = ~Petal.Length, color = ~Species, type = 'histogram',
        autobinx = TRUE, opacity = 0.7) %>%
  layout(title = "Distribution of Petal Length",
         xaxis = list(title = "Petal Length"),
         yaxis = list(title = "Count"))

```

### Sepal Length

```{r}
plot_ly(data = iris, x = ~Sepal.Length, color = ~Species, type = 'histogram',
        autobinx = TRUE, opacity = 0.7) %>%
  layout(title = "Distribution of Sepal Length",
         xaxis = list(title = "Sepal Length"),
         yaxis = list(title = "Count"))

```

## Density Plot {.tabset}

### Petal Width

```{r}


density_plot <- ggplot(iris, aes(x = Petal.Width, fill = Species)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
  labs(title = "Density Plot: Petal Width", x = "Petal Width", y = "Density")

ggplotly(density_plot)


```

### Sepal Width

```{r}


density_plot <- ggplot(iris, aes(x = Sepal.Width, fill = Species)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
  labs(title = "Density Plot: Sepal Width", x = "Sepal Width", y = "Density")

ggplotly(density_plot)


```

## Box Plot {.tabset}

### Petal Length

```{r}
plot_ly(data = iris, y = ~Petal.Length, color = ~Species, type = 'box') %>%
  layout(title = "Petal Length Across Species",
         yaxis = list(title = "Petal Length"))

```

### Sepal Length

```{r}
plot_ly(data = iris, y = ~Sepal.Length, color = ~Species, type = 'box') %>%
  layout(title = "Sepal Length Across Species",
         yaxis = list(title = "Sepal Length"))

```

# Predictor

```{r}

column(
  width = 4,
  sliderInput("sepal_length", "Sepal Length:", min = 4.1, max = 8, value = 5.5),
  sliderInput("sepal_width", "Sepal Width:", min = 2, max = 4.5, value = 3),
  sliderInput("petal_length", "Petal Length:", min = .1, max = 7, value = 4),
  sliderInput("petal_width", "Petal Width:", min = 0.1, max = 2.5, value = 1),
  actionButton("predict", "Predict Species")
)

renderText({
  input$predict
  
  isolate({
    new_data <- data.frame(
      Sepal.Length = input$sepal_length,
      Sepal.Width = input$sepal_width,
      Petal.Length = input$petal_length,
      Petal.Width = input$petal_width
    )
    

    distances <- apply(iris_kmeans$centers, 1, function(center) {
      sum((center - unlist(new_data))^2)
    })
    predicted_cluster <- which.min(distances)
    name_list <- list("setosa", "versicolor", "virginica")


    paste("Predicted Species Cluster:", name_list[[predicted_cluster]])
  })
})

```
